NumPy---Array Programming in Python

### 2020-02-13 authors.tex
Add script to generate auth/affil file

\author[1]{Charles R Harris}
\author[2,3,*]{K. Jarrod Millman}
\author[2,*]{St\'efan J. van der Walt}
\author[4,*]{Ralf Gommers}
\author[5]{Pauli Virtanen}
\author[6]{David Cournapeau}
\author[7]{Eric Wieser}
\author[8]{Julian Taylor}
\author[2]{Sebastian Berg}
\author[9]{Nathaniel J. Smith}
\author[10]{Robert Kern}
\author[2]{Matti Picus}
\author[11]{Stephan Hoyer}
\author[12]{Marten H. van Kerkwijk}
\author[13]{Matthew Brett}
\author[14]{Allan Haldane}
\author[15]{Jaime Fernandez del Rio}
\author[16]{Mark Wiebe}
\author[4,18,19]{Pearu Peterson}
\author[80]{David M. Cooke}
\author[21]{Pierre G\'erard-Marchant}
\author[22]{Kevin Sheppard}
\author[23]{Tyler Reddy}
\author[2]{Warren Weckesser}
\author[4]{Hameer Abbasi}
\author[24]{Christoph Gohlke}
\author[4]{Travis E. Oliphant}
\affil[1]{Independent Researcher, Logan, Utah, USA}
\affil[2]{Berkeley Institute for Data Science, University of California, Berkeley, Berkeley, CA, USA}
\affil[3]{Division of Biostatistics, University of California, Berkeley, Berkeley, CA, USA}
\affil[4]{Quansight LLC, Austin, TX, USA}
\affil[5]{Department of Physics and Nanoscience Center, University of Jyv\"askyl\"a, Jyv\"askyl\"a, Finland}
\affil[6]{Mercari JP, Tokyo, Japan}
\affil[7]{Department of Engineering, University of Cambridge, Cambridge, UK}
\affil[8]{Independent Researcher, Karlsruhe, Germany}
\affil[9]{Independent Researcher, Berkeley, CA, USA}
\affil[10]{Enthought, Inc., Austin, TX, USA}
\affil[11]{Google Research, Mountain View, CA, USA}
\affil[12]{Department of Astronomy \& Astrophysics, University of Toronto, Toronto, ON, Canada}
\affil[13]{School of Psychology, University of Birmingham, Edgbaston, Birmigham, UK}
\affil[14]{Department of Physics, Temple University, Philadelphia, PA, USA}
\affil[15]{Google, Zurich, Switzerland}
\affil[16]{Amazon, Seattle, Washington, USA}
\affil[18]{Independent Researcher, Saue, Estonia}
\affil[19]{Department of Mechanics and Applied Mathematics, Institute of Cybernetics at Tallinn Technical University, Tallinn, Estonia}
\affil[21]{France-IX Services, Paris, France}
\affil[22]{Department of Economics, University of Oxford, Oxford, UK}
\affil[23]{CCS-7, Los Alamos National Laboratory, Los Alamos, NM, USA}
\affil[24]{Laboratory for Fluorescence Dynamics, Biomedical Engineering Department, University of California, Irvine, Irvine, CA, USA}
\affil[80]{Not sure}
\affil[*]{numpy.articles@gmail.com}

>>>>

\author[1]{Charles R. Harris}
\author[2,3,*]{K. Jarrod Millman}
\author[2,*]{St\'efan J. van der Walt}
\author[4,*]{Ralf Gommers}
\author[5]{Pauli Virtanen}
\author[6]{David Cournapeau}
\author[7]{Eric Wieser}
\author[8]{Julian Taylor}
\author[2]{Sebastian Berg}
\author[9]{Nathaniel J. Smith}
\author[10]{Robert Kern}
\author[2]{Matti Picus}
\author[11]{Stephan Hoyer}
\author[12]{Marten H. van Kerkwijk}
\author[13]{Matthew Brett}
\author[14]{Allan Haldane}
\author[15]{Jaime Fernandez del Rio}
\author[16]{Mark Wiebe}
\author[4,17,18]{Pearu Peterson}
\author[19]{Pierre G\'erard-Marchant}
\author[20]{Kevin Sheppard}
\author[21]{Tyler Reddy}
\author[2]{Warren Weckesser}
\author[4]{Hameer Abbasi}
\author[22]{Christoph Gohlke}
\author[4]{Travis E. Oliphant}
\affil[1]{Independent Researcher, Logan, Utah, USA}
\affil[2]{Berkeley Institute for Data Science, University of California, Berkeley, Berkeley, CA, USA}
\affil[3]{Division of Biostatistics, University of California, Berkeley, Berkeley, CA, USA}
\affil[4]{Quansight LLC, Austin, TX, USA}
\affil[5]{Department of Physics and Nanoscience Center, University of Jyv\"askyl\"a, Jyv\"askyl\"a, Finland}
\affil[6]{Mercari JP, Tokyo, Japan}
\affil[7]{Department of Engineering, University of Cambridge, Cambridge, UK}
\affil[8]{Independent Researcher, Karlsruhe, Germany}
\affil[9]{Independent Researcher, Berkeley, CA, USA}
\affil[10]{Enthought, Inc., Austin, TX, USA}
\affil[11]{Google Research, Mountain View, CA, USA}
\affil[12]{Department of Astronomy \& Astrophysics, University of Toronto, Toronto, ON, Canada}
\affil[13]{School of Psychology, University of Birmingham, Edgbaston, Birmigham, UK}
\affil[14]{Department of Physics, Temple University, Philadelphia, PA, USA}
\affil[15]{Google, Zurich, Switzerland}
\affil[16]{Amazon, Seattle, Washington, USA}
\affil[17]{Independent Researcher, Saue, Estonia}
\affil[18]{Department of Mechanics and Applied Mathematics, Institute of Cybernetics at Tallinn Technical University, Tallinn, Estonia}
\affil[19]{France-IX Services, Paris, France}
\affil[20]{Department of Economics, University of Oxford, Oxford, UK}
\affil[21]{CCS-7, Los Alamos National Laboratory, Los Alamos, NM, USA}
\affil[22]{Laboratory for Fluorescence Dynamics, Biomedical Engineering Department, University of California, Irvine, Irvine, CA, USA}
\affil[*]{numpy.articles@gmail.com}

### 2020-02-13 body.tex
Add script to generate auth/affil file

Two Python array packages existed before NumPy.
The Numeric package began in the mid-1990s and provided an array object and array-aware functions
in Python, written in C, and linking to standard fast implementations of linear
algebra \cite{dubois1996numerical,Numericmanual}.
To handle large astronomical images coming from the Hubble Space Telescope,
a reimplementation of Numeric, called Numarray, added
support for structured arrays, flexible indexing, memory mapping, byte-order variants,
more efficient memory use, flexible IEEE error handling capabilities, and
better type casting rules \cite{greenfield2003numarray}.
While Numarray was highly compatible with Numeric, the two packages
had enough differences that it divided the community, until
2005, when NumPy emerged as a ``best of both worlds'' unification of Numeric
and Numarray \cite{oliphant2006guide}---combining Numarray's features with Numeric's
performance on small arrays and its rich C \emph{Application Programming Interface} (API).

@@

NumPy is a community of practice with a culture of
employing software engineering practice to improve collaboration and
reduce error \cite{millman2014developing}.  This culture is not only
adopted by leaders in the project but also enthusiastically taught to
newcomers. The NumPy team was early in adopting distributed revision
control and code review to improve collaboration on code, and
continuous testing that runs an extensive battery of automated tests for
every proposed change to NumPy.  The project has comprehensive,
high-quality documentation, integrated with the source
code \cite{vanderwalt2008scipy,harrington2008scipy,harrington2009scipy}.

@@

%For algorithms that do not fit the array programming approach,
%or require even faster implementations, a variety of tools exist (e.g., Cython, Numba, Pythran)
%that transparently accelerate bottlenecks.
%These tools allow programmers to express concepts using high-level array programming without sacrificing efficiency.
%In many cases, fast algorithms are already implemented by libraries in the scientific Python ecosystem;
%these primitives simplify writing code in pure Python.

% a brief word about accelerators perhaps?
% end with how this fits the bill for a huge number of users

@@

NumPy combines the expressive power of \emph{array programming}, 
the performance of C, and
the readability, usability, and versatility of Python in a mature,
well-tested, well-documented, and community-developed library.
%It builds on decades of work refining the array programming
%syntax from APL through Basis, Yorick, and others.
It also made it easy for other libraries to develop fast and
memory-efficient compiled code, usually in C or Fortran, that could manipulate
these arrays and pass them back to Python.

It was initially developed by students, faculty, and researchers to
provide an advanced, open-source array programming library for Python,
which was free to use and unencumbered by license servers, dongles, and the like.
There was a sense of building something consequential together,
for the benefit of many others.  Participating in
such an endeavor, within a welcoming community of like-minded
individuals, held a powerful attraction for many early contributors.

>>>>

Two Python array packages existed before NumPy.
The Numeric package began in the mid-1990s and provided an array object and array-aware functions
in Python, written in C, and linking to standard fast implementations of linear
algebra \cite{dubois1996numerical,Numericmanual}.
% FIXME steering physics codes at llnl
To handle large astronomical images coming from the Hubble Space Telescope,
a reimplementation of Numeric, called Numarray, added
support for structured arrays, flexible indexing, memory mapping, byte-order variants,
more efficient memory use, flexible IEEE error handling capabilities, and
better type casting rules \cite{greenfield2003numarray}.
While Numarray was highly compatible with Numeric, the two packages
had enough differences that it divided the community, until
2005, when NumPy emerged as a ``best of both worlds'' unification of Numeric
and Numarray \cite{oliphant2006guide}---combining Numarray's features with Numeric's
performance on small arrays and its rich C \emph{Application Programming Interface} (API).

@@

% this is the software engineering side of things
% as opposed to the interactive, exploration described above
NumPy is a community of practice with a culture of
employing software engineering practice to improve collaboration and
reduce error \cite{millman2014developing}.  This culture is not only
adopted by leaders in the project but also enthusiastically taught to
newcomers. The NumPy team was early in adopting distributed revision
control and code review to improve collaboration on code, and
continuous testing that runs an extensive battery of automated tests for
every proposed change to NumPy.  The project has comprehensive,
high-quality documentation, integrated with the source
code \cite{vanderwalt2008scipy,harrington2008scipy,harrington2009scipy}.

@@

@@

NumPy combines the expressive power of \emph{array programming}, 
the performance of C, and
the readability, usability, and versatility of Python in a mature,
well-tested, well-documented, and community-developed library.
%It builds on decades of work refining the array programming
%syntax from APL through Basis, Yorick, and others.
It also makes it easy for other libraries to develop fast and
memory-efficient compiled code, usually in C or Fortran, that could manipulate
these arrays and pass them back to Python.
For users that require faster implementations but do not want to write lower-level
compiled code, a variety of tools exist
(e.g., Cython, Numba, Pythran) that transparently accelerate bottlenecks.
These tools allow programmers to express concepts using high-level array
programming without sacrificing efficiency.
In many cases, fast algorithms are already implemented by libraries in the
scientific Python ecosystem; these primitives simplify writing code in pure
Python.

NumPy was initially developed by students, faculty, and researchers to
provide an advanced, open-source array programming library for Python,
which was free to use and unencumbered by license servers, dongles, and the like.
There was a sense of building something consequential together,
for the benefit of many others.  Participating in
such an endeavor, within a welcoming community of like-minded
individuals, held a powerful attraction for many early contributors.
